Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-05 p2/axioms.tex
Process todos
Clarifying inpedendence

\begin{mydef}[$\eqext$, $\strext$ and $\equiv$]
% Relation in the algebra
We write $A\eqext B$, or say that $B$ \emph{extends} $A$,
% intended interpretation:
to mean that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS\in\Dom{A}:\,A\aFS=B\aFS$.
% old inference rule: $ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$,
We can also see that $A\eqext B$ and $S\eqext T$ implies $A\cc S\eqext B\cc T$.
% symmetric:     aRb <=> bRa
% antisymmetric: aRb & bRa => a=b
% reflexive:     aRa
% transitive:    aRb & bRC => aRc
% preorder:      reflexive, transitive
% partial order: reflexive, transitive, antisymmetric
% equivalence:   reflexive, transitive, symmetric
% congruence:    https://en.wikipedia.org/wiki/Congruence_relation
% precongruence: congruence of a preorder
In other words, $\eqext$ is a preorder, and also a precongruence. % TODO

% Relation in the algebra
We write $A\equiv B$,
or say that $A$ and $B$ are \emph{equivalent,}
iff $A\eqext B$ and $B\eqext A$;
that is, $\equiv$ is the intersection of the preorder $\eqext$ with its inverse,
and so it is a congruence. % TODO
% old inference rule: $ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $.

@@

\Cref{rules_lemma} summarizes the basic properties of command pairs.
The claims listed are named \emph{\namecrefs{ax_separate_commute}}
as they can also be interpreted as inference rules in a pure
algebraic treatment of filesystem synchronization.
TODO elaborates this approach.
In \cref{axiom_proof} we offer proofs of each of the
\emph{\namecrefs{ax_separate_commute}} based on our model of filesystems.

@@

If $n=m$, then from \cref{ax_same_breaks} we must have $[y]=[z]$ and $[w]=[x]$
as otherwise the commands, in one order or the other, would break all filesystems.
The commands are not assertion commands, so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $y=w$.
% TODO Does this depend on |[D]|=1?
Therefore $[x]=[y]=[z]=[w]$ which with $y=w$ contradicts our assumption that the commands are different.

The remaining case is that $n\descendant m$ or $m\descendant n$.
If they are not directly related,
then from \cref{ax_distantrel_breaks} we know that
$\cxynv\cc\czwmv$ breaks all filesystems,
which contradicts the assumption that the commands are independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if the nodes are directly related, either
$\cxynv\cc\czwmv$ or $\czwmv\cc\cxynv$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

>>>>

\begin{mydef}[$\eqext$, $\strext$ and $\equiv$]
% Relation in the algebra
We write $A\eqext B$, or say that $B$ \emph{extends} $A$,
% intended interpretation:
to mean that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS\in\Dom{A}:\,A\aFS=B\aFS$.
% old inference rule: $ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$,
We can also see that $A\eqext B$ and $S\eqext T$ implies $A\cc S\eqext B\cc T$.
% symmetric:     aRb <=> bRa
% antisymmetric: aRb & bRa => a=b
% reflexive:     aRa
% transitive:    aRb & bRC => aRc
% preorder:      reflexive, transitive
% partial order: reflexive, transitive, antisymmetric
% equivalence:   reflexive, transitive, symmetric
% congruence:    https://en.wikipedia.org/wiki/Congruence_relation
% precongruence: congruence of a preorder
In other words, $\eqext$ is a preorder, and also a precongruence.

% Relation in the algebra
We write $A\equiv B$,
or say that $A$ and $B$ are \emph{equivalent,}
iff $A\eqext B$ and $B\eqext A$;
that is, $\equiv$ is the intersection of the preorder $\eqext$ with its inverse,
and so it is a congruence.
% old inference rule: $ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $.

@@

\Cref{rules_lemma} summarizes the basic properties of command pairs.
The claims listed are named \emph{\namecrefs{ax_separate_commute}}
as they can also be interpreted as inference rules in a pure
algebraic treatment of filesystem synchronization.
\Cref{sec_algebra} elaborates this approach.
In \cref{axiom_proof} we offer proofs of each of the
\emph{\namecrefs{ax_separate_commute}} based on our model of filesystems.

@@

If $n=m$, then from \cref{ax_same_breaks} we must have $[y]=[z]$ and $[w]=[x]$
as otherwise the commands, in one order or the other, would break all filesystems.
We also see that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $w=y$.
Therefore $n=m$, $[x]=[y]=[z]=[w]$ and $w=y$, which contradicts our assumption that the commands are different.

The remaining case is that $n\descendant m$ or $m\descendant n$.
If they are not directly related,
then from \cref{ax_distantrel_breaks} we know that
$\cxynv\cc\czwmv$ breaks all filesystems as they are not assertion commands,
which contradicts the assumption that the commands are independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if the nodes are directly related, either
$\cxynv\cc\czwmv$ or $\czwmv\cc\cxynv$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

### 2018-07-05 p2/works-inverse.tex
Clarifying inpedendence

\begin{mylem}\mlabel{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{mylem}
\begin{proof}
TODO: they are the same? FF?

In the following proof we reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
The proof is by contradiction;
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$(\cxynv\cc \czwmv)\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in two cases.
First, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[\cxynv\aFS(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Second, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of values at the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

>>>>

\begin{mylem}\mlabel{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{mylem}
\begin{proof}
TODO: they are the same? FF? assertion commands?

In the following proof we reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
The proof is by contradiction;
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$(\cxynv\cc \czwmv)\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.

We note that as the two commands are independent, they commute, and so
the proposition is symmetric.
Therefore, if either of the commands is an assertion command,
without loss of generality we can assume $\cxynv$ is an assertion command.
If so, 
$[\cxynv\aFS(m)]=[\FS(m)]$ even if $n=m$.
This means $\czwmv$ cannot break $\cxynv\aFS(m)$, which is a contradiction.

In the remaining cases neither command is an assertion command.

TODO continue here. Require not replacement?

Applying a command can only result in a broken filesystem in two cases.
First, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[\cxynv\aFS(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Second, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of values at the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}
